Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-05-02 second-draft/pnas-word-order-si.tex
WALS

@@

\section{UD Formalization}

\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    lifted\_case    &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
copula    &    NP    &    lifted\_cop    &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
auxiliary    &    VP    &    aux    &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
noun    &    genitive    &    nmod    &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
noun    &    relative clause    &    acl    &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
complementizer    &    S    &    lifted\_mark    &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
verb    &    PP    &    obl    &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
want    &    VP    &    xcomp    &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
verb    &    subject    &    nsubj    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
verb    &    manner adverb    &    advmod    &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
%	For clarity, we organize the correlations by the category of the `verb patterner'.
	For each correlation, we give our operationalization in terms of UD. %A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation. 
	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
\end{table*}

@@

@@

\tiny{
\begin{tabular}{|l|l|l|ll|l|l|}
	\hline
	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
% From Dryer	
acl  &  80   &   \textbf{85}$^{***}$   &   43   &   \textbf{82}$^{***}$   &   \textbf{71}$^{***}$  & $>50\%$  \citep{dryer1992greenbergian} \\
aux  &  12   &   \textbf{26}$^{***}$   &   \textbf{19}$^{***}$   &   49   &   \textbf{35}$^{**}$  & $<50\%$ \citep{dryer1992greenbergian} \\
lifted\_case  &  86   &   \textbf{81}$^{***}$   &   \textbf{40}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{67}$^{***}$ & $> 50 \%$ \citep{dryer1992greenbergian}  \\
lifted\_cop  &  94   &   \textbf{80}$^{***}$   &   59   &   \textbf{72}$^{***}$   &   \textbf{63}$^{**}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
lifted\_mark  &  76   &   \textbf{85}$^{***}$   &   \textbf{57}$^{*}$   &   \textbf{76}$^{***}$   &   \textbf{72}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
nmod  &  80   &   \textbf{82}$^{***}$   &   49   &   \textbf{71}$^{***}$   &   \textbf{68}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\ 
obl  &  88   &   \textbf{78}$^{***}$   &   \textbf{71}$^{***}$   &   46   &   \textbf{68}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
xcomp  &  88   &   \textbf{90}$^{***}$   &   \textbf{76}$^{***}$   &   \textbf{89}$^{***}$   &   \textbf{85}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
\hdashline
nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & $>50\%$ for manner adverbs; $\approx 50\%$ for negation, TAM, intensifiers \citep{dryer1992greenbergian} \\

\hline
% Other universals, not in Dryer
advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A vs. 83A]{wals} \\

\hline
% Typological data not available
appos  &  77   &   \textbf{77}$^{***}$   &   49   &   \textbf{67}$^{***}$   &   \textbf{62}$^{**}$ &  Unknown \\%($>50\%$? correlates with \textit{nmod} according to Greenberg U 23)  \\ % \cite{cinque2009greenberg} claims it correlates with OV/VO. However, it's not clear to me how closely appos matches the subject of the typological claim
lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & Unknown (cf. \cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & Unknown (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
iobj  &  74   &   49   &   \textbf{78}$^{***}$   &   \textbf{38}$^{*}$   &   62  & Unknown \\
vocative  &  46   &   \textbf{42}$^{*}$   &   59   &   45   &   \textbf{45}$^{*}$  & Unknown \\

\hline
% Uninterpretable
compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable (conflates articles and others) \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\ %TODO what else is conflated in det? why do we keep advmod but not det?
dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\

@@

\end{center}
\caption{Predictions on all UD relations occurring in more than one language. Numbers are the fraction of languages where the direction of the given relation agrees in direction with the \emph{obj} relation. In the last column, we indicate what direction would be expected typologically. ``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful. ``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
\label{tab:all-predictions}
\end{table*}

@@

\section{Results on Best Optimized Grammars}

In the main part, we reported the percentage of optimized grammars supporting correlations.
Here we report these numbers for the grammars that optimize each objective most successfully.
That is, for each language and objective, we chose among the eight optimized grammars the one that achieves the best value for the objective function on the validation partition of the treebanks -- that is, the one where the optimization procedure was most successful.

Numbers for these languages are shown in Table~\ref{table:results-dryer-best-only}.
%Confirming the results from the main paper, 
As expected, the correlations mostly have stronger prevalences than when considering all optimized languages.
For Dependency Length, the correlations that it predicts mostly have prevalences close to 100 \%, whereas both the UD languages and efficiency-optimized languages show somewhat lower percentages.
That is, dependency length mostly overshoots for the correlations that it predicts correctly.

\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    lifted\_case    &    86    &    98    &    69    &    88    &    83   \\
copula    &    NP    &    lifted\_cop    &    94    &    98    &    58    &    85    &    74   \\
auxiliary    &    VP    &    aux    &    88    &    84    &    90    &    61    &    73   \\
noun    &    genitive    &    nmod    &    80    &    98    &    75    &    88    &    91   \\
noun    &    relative clause    &    acl    &    80    &    100    &    69    &    88    &    90   \\
complementizer    &    S    &    lifted\_mark    &    76    &    100    &    76    &    82    &    87   \\
verb    &    PP    &    obl    &    88    &    98    &    88    &    75    &    91   \\
want    &    VP    &    xcomp    &    88    &    100    &    92    &    94    &    96   \\
verb    &    subject    &    nsubj    &    33    &    25    &    53    &    10    &    21   \\
verb    &    manner adverb    &    advmod    &    35    &    61    &    27    &    55    &    36   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
\end{tabular}
	\end{center}
	\caption{Percentages of grammars supporting correlations, among grammars that best optimize each objective on treebanks of a given language.}\label{table:results-dryer-best-only}
\end{table*}

@@

\section{Further Statistical Analyses}

In the main results table, we reported the fraction of optimized languages satisfying a correlation.
We show that essentially the same predicted fractions are obtained when controlling for the imbalance in distribution of language families.

To this end, we used the logistic mixed-effects regressions that were used to estimate significance to compute the expected prevalence of a correlation controlling for language families.
To obtain predictions controlling for languages and language families, we plot the logit-transformed posterior of $\alpha$ in Figure \ref{fig:posterior-prevalence}.

\begin{figure}
    \centering
    \includegraphics[scale=.7]{../results/correlations/figures/posterior_joint_with_ud.pdf}
	\caption{Posteriors of the prevalences of the correlations among languages optimized for Predictability (red), parseability (green), and efficiency (blue), controlling for languages and language families, with prevalence in the UD languages.}
    \label{fig:posterior-prevalence}
\end{figure}

\begin{figure}
    \centering
%     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
 %    \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
  %   \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
   %  \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
     \includegraphics[scale=.45]{../results/correlations/figures/posterior-satisfied-universals-together.pdf}

	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies. Efficiency predicts all ten correlations with high posterior probability.}
    \label{fig:posterior-satisfied-count}
\end{figure}

We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
We conducted the same analysis for predictability, parseability, and dependency length.
We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$. The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.

%\begin{figure}
%    \centering
%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-efficiency.pdf}
%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-langmod.pdf}
%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-parseability.pdf}
%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-depl.pdf}
%	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
%    \label{fig:posterior-satisfied-count}
%\end{figure}

%We argued that efficiency predicts all 10 correlation universals in the direction manifested in the UD corpora, as for each universal, the posterior probability that less than $50 \%$ of the optimized grammars satisfy the correlation was less than $0.05$.
%To address potential multiple comparisons problems, we use a single logistic model to show that efficiency predicts all ten correlations with high posterior probability.

%For each objective function, we constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency, with random effects for language and language family, correlated across the ten dependencies.

%We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
%The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
%The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$. The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.

@@

@@

@@

\section{Implementation Details}

\paragraph{Locality}
In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.

\paragraph{Language Model}
We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
Given the small size of the corpora, this limit is only attained only for few languages.
In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
The model predicts both the next word and its POS tags in each step.
Using POS tags is intended to prevent overfitting on small corpora.
These decisions were made before evaluating word order properties.

\paragraph{Parser}
We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}. This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work. The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word. This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.

To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags. Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged. Parsing from POS tags reduces early overfitting.
This decision was made before evaluating word order properties.

\paragraph{Hyperparameters}
For locality, predictability, and parseability, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function on the language. We then used this setting for creating optimized word order grammars. % TODO more detail and maybe an example here
% TODO all design changes were made before examining word order universal results

% TODO Discuss iconicity somewhere somehow

All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
We do not use pretrained embeddings \citep{peters2018deep}: these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.

@@

We update word order grammar parameters $\theta$ using SGD with momentum.
For the language model parameters $\phi$, we use plain SGD. 
For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.

@@

Recall that, for each language, we created 8 optimized languages for each optimization criterion.
We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
This maximizes statistical power for detecting and quantifying correlations between the \textit{obj} relation and other relations.

@@

\subsection{CYK Parsers}

@@

@@

\section{}

\bibliographystyle{acl_natbib}
\bibliography{everything}

>>>>

\usepackage{longtable}

\usepackage{natbib}

@@

\section{Formalization of Correlations}

Here we describe how we selected the correlations in Table 1 of the main paper, and how we formalized these using syntactic relations defined by Universal Dependencies.

We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding recent objections by~\cite{dunn2011evolved}, see follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
While most of the other relations also correspond to UD relations, some are not annotated reliably.
We were able formalize eleven out of Dryer's seventeen correlations in UD. Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations.

We excluded four correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
All three types of elements occur at most in parts of the 51 UD languages, and none of them is annotated reliably in those languages where they occur.
Among these three types of elements, the one most prominent in our sample of 51 languages is articles (occurring in many European languages); UD subsumes them under the \emph{det} relation, which is also used for other frequent elements, such as demonstratives and quantifiers.

We also excluded the verb-manner adverb correlation.
UD does not distinguish manner adverbs from other elements labeled as adverbs, such as sentence-level adverbs and negation markers, whose ordering is very different from manner adverbs.
All types of adverbs are unified under the \emph{advmod} relation.
In the real orderings in our sample of 51 UD languages, the dominant ordering of \emph{advmod} almost always matches that of subjects -- that is, \emph{advmod} dependents are ordered after the verb only in VSO languages.
This ordering behavior is very different from that documented for manner adverbs by Dryer.

We further excluded the verb-subject correlation, which is not satisfied by much more than half of the world's languages (51 \% among those with annotation in \cite{wals-81}, with clear violation in 35.4 \%).
It is satisfied only in 33\% of our sample of 51 UD languages, as quantified using the grammars we extracted.
Dryer \cite{dryer1992greenbergian} counts this as a correlation since he describes the distribution of subject order as an interaction between a weak correlation with object order, and a very strong dominance principle favoring SV/SO orderings.
We focus on the modeling of correlations, and such dominance principles are beyond the scope of our experiments.
We therefore excluded this correlation.
%We note that efficiency makes predictions compatible with the pattern observed in our sample of 51 languages, potentially in line with the noisy-channel explanation of SVO described by~\cite{gibson2013noisy} (see Table~\ref{tab:all-predictions}).

TODO collapsing

The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.

\begin{table*}[ht]
	\begin{center}
\begin{tabular}{|ll|l|l}
	\hline
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       \\ 
	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    lifted\_case   & G3, G4   \\ \hline
copula  verb  &    predicate    &    lifted\_cop   & --    \\\hline
`want'    &    VP    &    xcomp   & G15   \\\hline
tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & G16, G13  \\
negative auxiliary    &    VP    &     \\ \hline
complementizer    &    S    &   \multirow{2}{*}{lifted\_mark}  & --    \\
adverbial subordinator & S & \\ \hline
noun    &    genitive    &    nmod & G2, G23   \\ \hline
noun    &    relative clause    &    acl   &  G24    \\ \hline
adjective & standard of comparison & \multirow{2}{*}{obl}\\
verb    &    PP    &    & G22   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{3}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations according to Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies. We also provide the corresponding numbers in Greenberg's original study.}\label{table:greenberg-dryer}
\end{table*}

%croft2011greenbergian
%dryer2011evidence

\section{Supporting Analyses for Study 1}

\paragraph{Other tests, maybe som viz}

\paragraph{Pareto Curve}

It has to be convex, so don't need to do more.

@@

We also report results for grammars optimized for predictability and parseability individually.

Further, we report results for grammars optimized for dependency length minimization (DLM).
DLM can be seen as a heuristic approximation to efficiency.
We also found that DLM is predicted by efficiency optimization.
Indeed, we show that optimizing for DLM makes predictions similar to efficiency optimization.

@@

\small{
\begin{tabular}{|l|l|l|ll|l|l|}
	\hline
	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
% From Dryer	
acl  &  80   &   \textbf{85}$^{***}$   &   43   &   \textbf{82}$^{***}$   &   \textbf{71}$^{***}$  & $>50\%$  \citep{dryer1992greenbergian} \\
aux  &  12   &   \textbf{26}$^{***}$   &   \textbf{19}$^{***}$   &   49   &   \textbf{35}$^{**}$  & $<50\%$ \citep{dryer1992greenbergian} \\
lifted\_case  &  86   &   \textbf{81}$^{***}$   &   \textbf{40}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{67}$^{***}$ & $> 50 \%$ \citep{dryer1992greenbergian}  \\
lifted\_cop  &  94   &   \textbf{80}$^{***}$   &   59   &   \textbf{72}$^{***}$   &   \textbf{63}$^{**}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
lifted\_mark  &  76   &   \textbf{85}$^{***}$   &   \textbf{57}$^{*}$   &   \textbf{76}$^{***}$   &   \textbf{72}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
nmod  &  80   &   \textbf{82}$^{***}$   &   49   &   \textbf{71}$^{***}$   &   \textbf{68}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\ 
obl  &  88   &   \textbf{78}$^{***}$   &   \textbf{71}$^{***}$   &   46   &   \textbf{68}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
xcomp  &  88   &   \textbf{90}$^{***}$   &   \textbf{76}$^{***}$   &   \textbf{89}$^{***}$   &   \textbf{85}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
\hdashline
nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\

\hline
% Other universals, not in Dryer
advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & $>50\%$ (cf. \cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A, 83A]{wals} \\ %

\hline
% Typological data not available
appos  &  77   &   \textbf{77}$^{***}$   &   49   &   \textbf{67}$^{***}$   &   \textbf{62}$^{**}$ &  Unknown \\%($>50\%$? correlates with \textit{nmod} according to Greenberg U 23)  \\ % \cite{cinque2009greenberg} claims it correlates with OV/VO. However, it's not clear to me how closely appos matches the subject of the typological claim
lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
iobj  &  74   &   49   &   \textbf{78}$^{***}$   &   \textbf{38}$^{*}$   &   62  & Unknown \\
vocative  &  46   &   \textbf{42}$^{*}$   &   59   &   45   &   \textbf{45}$^{*}$  & Unknown \\

\hline
% Uninterpretable
compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\
dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\
advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & Uninterpretable \\

@@

\end{center}
\caption{Predictions on all UD relations occurring in more than one language. Numbers are the fraction of languages where the direction of the given relation agrees in direction with the \emph{obj} relation. In the last column, we indicate what direction would be expected typologically. The first section correspond to universals from Dryer. ``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful. ``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
\label{tab:all-predictions}
\end{table*}

@@

@@

@@

The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).

@@

\subsection{Differentiable Ordering Grammars}

In order to carry out optimization of grammars, we extend the space of grammars to include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
This makes efficiency a differentiable function of the grammar parameters, and enables optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.

The parameters of such a differentiable word order grammar are as follows. 
For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p. 184) without replacement.

If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.

\subsection{Extracting Grammars from Datasets}
We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.

\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}

\textbf{rewrite this}

@@

%We extract grammars for the actual languages using a Bayesian data analysis method.
%We specify a simple Bayesian prior over differentiable ordering grammars:
%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
%for each $\tau$.
%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
%We extract

\section{Neural Network Architectures}

In this section, we describe the details of the neural network architectures.
All choices described in this section were fixed before evaluating word order properties and the efficiency of real language.

%\paragraph{Locality}
%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.

\paragraph{Estimating Predictability}
We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
Given the small size of the corpora, this limit is only attained only for few languages.
In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
The model predicts both the next word and its POS tags in each step.
Using POS tags is intended to prevent overfitting on small corpora.
These decisions were made before evaluating word order properties.

\paragraph{Estimating Parseability}
We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}. This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work. The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word. This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.

To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags. Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
This decision was made before evaluating word order properties.

\paragraph{Hyperparameters}

\textbf{TODO study 1 only has 1 set of parameters}

Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (predictability or parseability) on the language. We then used this setting for creating optimized word order grammars.

\textbf{TODO random seeds for efficiency}

All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.

@@

We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
For the language model parameters $\phi$, we use plain SGD. 
For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.

@@

Recall that, for each language, we created 8 optimized languages for each optimization criterion.
We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.

@@

\subsection{CKY Parsers}

@@

\textbf{do same for language model, and for the other random grammars}

@@

\section{Study 2, Correlation between Universals and Efficiency}
